---
title: "Regression"
author: "Priyank Thakkar"
date: "25/10/2021"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

```{r Library}
library(tidyverse)
library(corrplot)
library(MASS)
library(car)
```

# Dataset Format : SWISS

Here, from previous project "[R]-PCA", we have selected Swiss data set as our dataset to perform Linear Regression.
This data set is about Standardized fertility measure and socio-economic indicators for each of 47 French-speaking provinces of Switzerland at about 1888. Each of which is in percent, i.e., in [0, 100].
[,1]	Fertility	Ig, ‘common standardized fertility measure’
[,2]	Agriculture	% of males involved in agriculture as occupation
[,3]	Examination	% draftees receiving highest mark on army examination
[,4]	Education	% education beyond primary school for draftees.
[,5]	Catholic	% ‘catholic’ (as opposed to ‘protestant’).
[,6]	Infant.Mortality	live births who live less than 1 year.

All variables but ‘Fertility’ give proportions of the population.

```{r summary, exercise=TRUE}
data(swiss)
summary(swiss)
```
# Variable Distribution

```{r Var_dist, exercise=TRUE}
pairs(swiss, panel = panel.smooth, main = "swiss data", col = 3 + (swiss$Catholic > 50))
```

# ANOVA

**Analysis of Variance** technique partitions the total variability in the sample data into component parts - variance explained by the regression line and the residual variance unexplained by the regression line. 

ANOVA F-Test
$F_0 = MS_R/MS_E$
Under null hypothesis $F_0$ has F-distribution with $(a-1)$ and $a(n-1)$ degrees of freedom

Lets perform ANOVA on our model from previous section   

```{r boxplot_aov, exercise = TRUE}
boxplot(swiss, xlab="Six Variables", ylab="Percentage", main="Box Plot of Variables")
```
### Findings/Conclusion (Boxplot)

* Catholic variable covers wide range of values
* Infant.Mortality variable is very condensed
* Education and Fertility seems to have some outliers

```{r aov_example_, exercise=TRUE}
fit = aov(Fertility~ ., data=swiss)
anova(fit)
```
# Correlation among Input Variables
Are the input variables correlated? Lets start with computing and visualizing the correlation matrix

```{r corrplot_dist, exercise=TRUE}
cor(swiss)
corrplot.mixed(cor(swiss[,-1]), order="hclust", tl.col="black")

pairs(swiss)
```
### Findings/Conclusion (Correlation)

* All correlations with Fertility are less than **0.7**, indicating no signs of strong multicollinearity.
* Correlations are between **0.3-0.7**, indicating mild multicollinearity.
* Plot shows linear relationship between **Agriculture and Examination**. Moreover, also between
**Examination and Education**.

# Regression 

Building empirical models

```{r check_dist_, exercise=TRUE}
ggplot(data = swiss, aes(Fertility)) + 
  geom_histogram(mapping = aes(x = Fertility, y = stat(density)), bins = 7) +
   stat_function(
    fun = dnorm, 
    args = list(mean = mean(swiss$Fertility), sd = sd(swiss$Fertility)), 
    lwd = 2, 
    col = 'red')

Fertility <- ggplot(data = swiss, aes(sample = Fertility))

Fertility + 
  stat_qq(distribution = stats::qnorm) + stat_qq_line() +
  labs(y = 'sample quantiles', x = 'theoretical quantiles') + 
  theme(text = element_text(size = 16))
```
### Findings/Conclusion (Histogram)

* In Histogram Plot is skewed or leaned towards the right side a little bit.
* Fertility rates are mostly between 60-90%


## Fit a Model 

```{r lm_example, exercise = TRUE}
model1 <- lm(Fertility ~ ., swiss)
summary(model1)
```
## Residual Analysis

```{r stepwise}
# stepAIC(model1, direction = "both")
# link of reference : https://ashutoshtr.medium.com/what-is-stepaic-in-r-a65b71c9eeba

model2 <- step(model1)
```
* The final AIC of model Achieved = 189.86
* Alternatively use this command for to check the AIC values available.
* And select the model with having the lowest AIC number, means Small lose of data while performing that model.

```{r summary model2}
summary(model2)
```

```{r avon}
anova(model1, model2)
# drop1(model1,test='F')
```
### Findings/Conclusion (Model1 and Model2)

* By both a T-test and an ANOVA F test we find **Examination** does not have significant effect on **Fertility**.
* Final Model2 $(Fertility ~ (Agriculture + Education + Catholic + Infant.Mortality), data = swiss)$ 

```{r residual plot}
par(mfrow=c(2,2))
plot(model2)
```

## Other Diagnostic functions 
Other functions you can try:

* coefficients(model2) # model coefficients 
* confint(model2, level=0.95) # CIs for model parameters
* fitted(model2) # predicted values
* residuals(model2) # residuals
* anova(model2) # anova table
* vcov(model2) # covariance matrix for model parameters
* influence(model2) # regression diagnostics

```{r reg_func, exercise = TRUE, exercise.setup = "reg_example"}
confint(model2, level=0.95)
```

```{r check_res, eval=FALSE, include=FALSE}
# diagnostic plots
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page
plot(model2)
```

## Additional Regression Diagnostics
```{r outlier_lev, exercise =TRUE, exercise.setup = "reg_example"}
# Assessing Outliers
outlierTest(model2) # Bonferonni p-value for most extreme obs
qqPlot(model2, main="QQ Plot") #qq plot for studentized resid
leveragePlots(model2) # leverage plots
```