---
title: "Untitled"
author: "shilpa gupta"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ANOVA

**Analysis of Variance** technique partitions the total variability in the sample data into component parts - variance explained by the regression line and the residual variance unexplained by the regression line. 

ANOVA F-Test
$F_0 = MS_R/MS_E$
Under null hypothesis $F_0$ has F-distribution with $(a-1)$ and $a(n-1)$ degrees of freedom

Lets perform ANOVA on our model from previous section   

```{r boxplot_aov, exercise = TRUE}
boxplot(mpg~as.factor(cyl),data=mtcars, main="Car Mileage Data",
   xlab="Number of Cylinders", ylab="Miles Per Gallon")
```


```{r aov_example_, exercise=TRUE}
fit = aov(mpg~as.factor(cyl), data=mtcars)
anova(fit)
```

## Correlation among Input Variables
Are the input variables correlated? Lets start with computing and visualizing the correlation matrix
```{r pca,exercise=TRUE}
library(corrplot)
library(tidyverse)
library(car)
corrplot.mixed(cor(mtcars[,-1]), order="hclust", tl.col="black")
```

## Regression
Building empirical models

Before we begin regression modeling, we need to check the distribution of the response we are trying to predict

```{r dist_extra, eval=FALSE, include=FALSE}
#qqnorm(mtcars$mpg)
#qqline(mtcars$mpg, distribution = qnorm )
```

```{r check_dist_, exercise=TRUE}
ggplot(data = mtcars, aes(mpg)) + geom_histogram(mapping = aes(x = mpg, y = stat(density)), bins = 7) +
   stat_function(
    fun = dnorm, 
    args = list(mean = mean(mtcars$mpg), sd = sd(mtcars$mpg)), 
    lwd = 2, 
    col = 'red')
mpg <- ggplot(data = mtcars, aes(sample = mpg))
mpg + stat_qq(distribution = stats::qnorm) + stat_qq_line() +
  labs(y = 'sample quantiles', x = 'theoretical quantiles') + 
  theme(text = element_text(size = 16))
```


Some of the questions we are trying to address with performing regression analysis
Q: Is the model statistically different from zero (p-value < 0.05)?
Q: Is the model predictive (Adjusted R-squared > 0.95).

```{r reg_model_, eval=FALSE, include=FALSE}
full_model = "cyl+disp+hp+drat+wt+qsec+vs+am+gear+carb" 
sig_model = "am+qsec+wt"
```


Fit a full model
```{r reg_example, exercise=TRUE}
fit = lm(mpg~wt+ as.factor(cyl), data=mtcars)
summary(fit)
```

### Residual Analysis
```{r reg_resi, exercise =TRUE, exercise.setup = "reg_example"}
plot(fit)
```


## Other Diagnostic functions 
Other functions you can try:

* coefficients(fit) # model coefficients 
* confint(fit, level=0.95) # CIs for model parameters
* fitted(fit) # predicted values
* residuals(fit) # residuals
* anova(fit) # anova table
* vcov(fit) # covariance matrix for model parameters
* influence(fit) # regression diagnostics

```{r reg_func, exercise = TRUE, exercise.setup = "reg_example"}
confint(fit, level=0.95)
```

```{r check_res, eval=FALSE, include=FALSE}
# diagnostic plots
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page
plot(fit)
```

## Additional Regression Diagnostics
```{r outlier_lev, exercise =TRUE, exercise.setup = "reg_example"}
# Assessing Outliers
outlierTest(fit) # Bonferonni p-value for most extreme obs
qqPlot(fit, main="QQ Plot") #qq plot for studentized resid
leveragePlots(fit) # leverage plots
```